{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-11-30 ident_manu.Rnw
first paragraph of ident results done

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in \Sexpr{locO2$tot} of the \Sexpr{length(p1z1pars)} (\Sexpr{round(100 * locO2$tot/length(p1z1pars), 0)}\% of total) evaluated parameters in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to \Sexpr{locO2$bycat$Optics} parameters for optics (\Sexpr{locO2$lncat$Optics} of all optic parameters), \Sexpr{locO2$bycat$`Organic Matter`} for organic matter (\Sexpr{locO2$lncat$`Organic Matter`}), \Sexpr{locO2$bycat$Phytoplankton} for phytoplankton (\Sexpr{locO2$lncat$Phytoplankton}), \Sexpr{locO2$bycat$Temperature} for temperature (\Sexpr{locO2$lncat$Temperature}), and \Sexpr{locO2$bycat$Zooplankton} for zooplankton (\Sexpr{locO2$lncat$Zooplankton}). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ \Sexpr{locO2$sensrng$error[1]} for \Sexpr{locO2$sensrng$Parameter[1]} (\Sexpr{tolower(locO2$sensrng$Category[1])}) to \Sexpr{locO2$sensrng$error[2]} for \Sexpr{locO2$sensrng$Parameter[2]} (\Sexpr{tolower(locO2$sensrng$Category[2])}), whereas average sensitivity among all parameters was $L1 = $ \Sexpr{locO2$sensave}. Within categories (excluding temperature), sensitivity ranged from \Sexpr{locO2$sensrngcat$Optics$error[1]} (\Sexpr{locO2$sensrngcat$Optics$Parameter[1]}) to \Sexpr{locO2$sensrngcat$Optics$error[2]} (\Sexpr{locO2$sensrngcat$Optics$Parameter[2]}) for optics, \Sexpr{locO2$sensrngcat$`Organic Matter`$error[1]} (\Sexpr{locO2$sensrngcat$`Organic Matter`$Parameter[1]}) to \Sexpr{locO2$sensrngcat$`Organic Matter`$error[2]} (\Sexpr{locO2$sensrngcat$`Organic Matter`$Parameter[2]}) for organic matter, \Sexpr{locO2$sensrngcat$Phytoplankton$error[1]} (\Sexpr{locO2$sensrngcat$Phytoplankton$Parameter[1]}) to \Sexpr{locO2$sensrngcat$Phytoplankton$error[2]} (\Sexpr{locO2$sensrngcat$Phytoplankton$Parameter[2]}) for phytoplankton, and \Sexpr{locO2$sensrngcat$Zooplankton$error[1]} (\Sexpr{locO2$sensrngcat$Zooplankton$Parameter[1]}) to \Sexpr{locO2$sensrngcat$Zooplankton$error[2]} (\Sexpr{locO2$sensrngcat$Zooplankton$Parameter[2]}) for zooplankton.  Average sensitivity values in each category were $L1 = $ \Sexpr{locO2$sensavecat$Optics} for optics, \Sexpr{locO2$sensavecat$`Organic Matter`} for organic matter, \Sexpr{locO2$sensavecat$Temperature} for temperature, \Sexpr{locO2$sensavecat$Phytoplankton} for phytoplankton, and \Sexpr{locO2$sensavecat$Zooplankton} for zooplankton.

@@

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All additional variables were sensitive to the same parameters (\Sexpr{locO2$tot} of \Sexpr{length(p1z1pars)} evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from \Sexpr{scinot(aveall[2])} for \Sexpr{names(aveall)[2]} (\cref{tab:irrsens}) to \Sexpr{scinot(aveall[length(aveall)])} for \Sexpr{names(aveall)[length(aveall)]} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = \Sexpr{scinot(aveall[1])}).  For each variable, $L1$ ranged from \Sexpr{locNH4$sensrng[1, 2]} (\Sexpr{locNH4$sensrng[1, 1]}) to \Sexpr{locNH4$sensrng[2, 2]} (\Sexpr{locNH4$sensrng[2, 1]}) for ammonium (\cref{tab:nh4sens}), \Sexpr{locChla_mg_tot$sensrng[1, 2]} (\Sexpr{locChla_mg_tot$sensrng[1, 1]}) to \Sexpr{locChla_mg_tot$sensrng[2, 2]} (\Sexpr{locChla_mg_tot$sensrng[2, 1]}) for \ac{chla} (\cref{tab:chlsens}), \Sexpr{locirradiance$sensrng[1, 2]} (\Sexpr{locirradiance$sensrng[1, 1]}) to \Sexpr{locirradiance$sensrng[2, 2]} (\Sexpr{locirradiance$sensrng[2, 1]}) for irradiance (\cref{tab:irrsens}), \Sexpr{locNO3$sensrng[1, 2]} (\Sexpr{locNO3$sensrng[1, 1]}) to \Sexpr{locNO3$sensrng[2, 2]} (\Sexpr{locNO3$sensrng[2, 1]}) for nitrate (\cref{tab:no3sens}), \Sexpr{locOM1$sensrng[1, 2]} (\Sexpr{locOM1$sensrng[1, 1]}) to \Sexpr{locOM1$sensrng[2, 2]} (\Sexpr{locOM1$sensrng[2, 1]}) for \ac{pom} (\cref{tab:om1sens}),  \Sexpr{locOM2$sensrng[1, 2]} (\Sexpr{locOM2$sensrng[1, 1]}) to \Sexpr{locOM2$sensrng[2, 2]} (\Sexpr{locOM2$sensrng[2, 1]}) for \ac{dom} (\cref{tab:om2sens}), and \Sexpr{locPO4$sensrng[1, 2]} (\Sexpr{locPO4$sensrng[1, 1]}) to \Sexpr{locPO4$sensrng[2, 2]} (\Sexpr{locPO4$sensrng[2, 1]}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to \Sexpr{names(maxcat$NH4)} parameters (average $L1$ = \Sexpr{maxcat$NH4} across all parameters in the category), \ac{chla} to \Sexpr{names(maxcat$Chla_mg_tot)} ($L1$ = \Sexpr{maxcat$Chla_mg_tat}), irradiance to \Sexpr{names(maxcat$irradiance)} ($L1$ = \Sexpr{maxcat$irradiance}), nitrate to \Sexpr{names(maxcat$NO3)} ($L1$ = \Sexpr{maxcat$NO3}), \ac{pom} to \Sexpr{names(maxcat$OM1)} ($L1$ = \Sexpr{maxcat$OM1}), \ac{dom} to \Sexpr{names(maxcat$OM2)} ($L1$ = \Sexpr{maxcat$OM2}), and phosphate to \Sexpr{names(maxcat$PO4)} ($L1$ = \Sexpr{maxcat$PO4}).  Finally, average sensititivy between parameter categories independent of the state variables ranged from \Sexpr{avecat[1]} for \Sexpr{names(avecat)[1]} (average $L1$ across all variables) to \Sexpr{avecat[length(avecat)]} for \Sexpr{names(avecat)[length(avecat)]}.

@@

%' data(cat_ident)
%' data(top_ident)
%' collcatpers <- lapply(cat_ident, function(x){
%' colltoppers <- lapply(top_ident, function(x){
%' maxtempcoll <- scinot(max(cat_ident$Temperature$coll))
%' The identifiability analyses suggested that most parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Parameter identifiability (as $\gamma$) increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identbox}).  By category, identifiability was lowest for parameter subsets in the phytoplankton (\Sexpr{collcatpers$Phytoplankton[1]}\% less than $\gamma = 15$, \Sexpr{collcatpers$Phytoplankton[2]}\% less than $\gamma = 10$)  and zooplankton categories (\Sexpr{collcatpers$Zooplankton[1]}\%, less than $\gamma = 15$, \Sexpr{collcatpers$Zooplankton[2]}\%, less than $\gamma = 10$), whereas a majority of combinations for temperature were identifiable (\Sexpr{collcatpers$Temperature[1]}\% less than $\gamma = 15$, \Sexpr{collcatpers$Temperature[2]}\%, less than $\gamma = 10$).  All subset combinations for optics and organic matter parameters had $\gamma < 10$.  All parameter subsets for choosing the top, top two, and top three parameters in each category were identifiable (\cref{fig:identbox}), whereas a majority were identifiable for choosing the top four (\Sexpr{colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{colltoppers$top4[2]}\% less than $\gamma = 10$) and top five (\Sexpr{colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{colltoppers$top4[2]}\%, less than $\gamma = 10$) parameters.  

>>>>

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in \Sexpr{locO2$tot} of the \Sexpr{length(p1z1pars)} (\Sexpr{round(100 * locO2$tot/length(p1z1pars), 0)}\% of total) evaluated parameters in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to \Sexpr{locO2$bycat$Optics} parameters for optics (\Sexpr{locO2$lncat$Optics} of all optic parameters), \Sexpr{locO2$bycat$`Organic Matter`} for organic matter (\Sexpr{locO2$lncat$`Organic Matter`}), \Sexpr{locO2$bycat$Phytoplankton} for phytoplankton (\Sexpr{locO2$lncat$Phytoplankton}), \Sexpr{locO2$bycat$Temperature} for temperature (\Sexpr{locO2$lncat$Temperature}), and \Sexpr{locO2$bycat$Zooplankton} for zooplankton (\Sexpr{locO2$lncat$Zooplankton}). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ \Sexpr{locO2$sensrng$error[1]} for \Sexpr{locO2$sensrng$Parameter[1]} (\Sexpr{tolower(locO2$sensrng$Category[1])}) to \Sexpr{locO2$sensrng$error[2]} for \Sexpr{locO2$sensrng$Parameter[2]} (\Sexpr{tolower(locO2$sensrng$Category[2])}), whereas average sensitivity among all parameters was $L1 = $ \Sexpr{locO2$sensave}. Within categories (excluding temperature), sensitivity ranged from \Sexpr{locO2$sensrngcat$Optics$error[1]} (\Sexpr{locO2$sensrngcat$Optics$Parameter[1]}) to \Sexpr{locO2$sensrngcat$Optics$error[2]} (\Sexpr{locO2$sensrngcat$Optics$Parameter[2]}) for optics, \Sexpr{locO2$sensrngcat$`Organic Matter`$error[1]} (\Sexpr{locO2$sensrngcat$`Organic Matter`$Parameter[1]}) to \Sexpr{locO2$sensrngcat$`Organic Matter`$error[2]} (\Sexpr{locO2$sensrngcat$`Organic Matter`$Parameter[2]}) for organic matter, \Sexpr{locO2$sensrngcat$Phytoplankton$error[1]} (\Sexpr{locO2$sensrngcat$Phytoplankton$Parameter[1]}) to \Sexpr{locO2$sensrngcat$Phytoplankton$error[2]} (\Sexpr{locO2$sensrngcat$Phytoplankton$Parameter[2]}) for phytoplankton, and \Sexpr{locO2$sensrngcat$Zooplankton$error[1]} (\Sexpr{locO2$sensrngcat$Zooplankton$Parameter[1]}) to \Sexpr{locO2$sensrngcat$Zooplankton$error[2]} (\Sexpr{locO2$sensrngcat$Zooplankton$Parameter[2]}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ \Sexpr{locO2$sensavecat$Optics} for optics, \Sexpr{locO2$sensavecat$`Organic Matter`} for organic matter, \Sexpr{locO2$sensavecat$Temperature} for temperature, \Sexpr{locO2$sensavecat$Phytoplankton} for phytoplankton, and \Sexpr{locO2$sensavecat$Zooplankton} for zooplankton.

@@

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All additional variables were sensitive to the same parameters (\Sexpr{locO2$tot} of \Sexpr{length(p1z1pars)} evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from \Sexpr{scinot(aveall[2])} for \Sexpr{names(aveall)[2]} (\cref{tab:irrsens}) to \Sexpr{scinot(aveall[length(aveall)])} for \Sexpr{names(aveall)[length(aveall)]} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = \Sexpr{scinot(aveall[1])}).  For each variable, $L1$ ranged from \Sexpr{locNH4$sensrng[1, 2]} (\Sexpr{locNH4$sensrng[1, 1]}) to \Sexpr{locNH4$sensrng[2, 2]} (\Sexpr{locNH4$sensrng[2, 1]}) for ammonium (\cref{tab:nh4sens}), \Sexpr{locChla_mg_tot$sensrng[1, 2]} (\Sexpr{locChla_mg_tot$sensrng[1, 1]}) to \Sexpr{locChla_mg_tot$sensrng[2, 2]} (\Sexpr{locChla_mg_tot$sensrng[2, 1]}) for \ac{chla} (\cref{tab:chlsens}), \Sexpr{locirradiance$sensrng[1, 2]} (\Sexpr{locirradiance$sensrng[1, 1]}) to \Sexpr{locirradiance$sensrng[2, 2]} (\Sexpr{locirradiance$sensrng[2, 1]}) for irradiance (\cref{tab:irrsens}), \Sexpr{locNO3$sensrng[1, 2]} (\Sexpr{locNO3$sensrng[1, 1]}) to \Sexpr{locNO3$sensrng[2, 2]} (\Sexpr{locNO3$sensrng[2, 1]}) for nitrate (\cref{tab:no3sens}), \Sexpr{locOM1$sensrng[1, 2]} (\Sexpr{locOM1$sensrng[1, 1]}) to \Sexpr{locOM1$sensrng[2, 2]} (\Sexpr{locOM1$sensrng[2, 1]}) for \ac{pom} (\cref{tab:om1sens}),  \Sexpr{locOM2$sensrng[1, 2]} (\Sexpr{locOM2$sensrng[1, 1]}) to \Sexpr{locOM2$sensrng[2, 2]} (\Sexpr{locOM2$sensrng[2, 1]}) for \ac{dom} (\cref{tab:om2sens}), and \Sexpr{locPO4$sensrng[1, 2]} (\Sexpr{locPO4$sensrng[1, 1]}) to \Sexpr{locPO4$sensrng[2, 2]} (\Sexpr{locPO4$sensrng[2, 1]}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to \Sexpr{names(maxcat$NH4)} parameters (average $L1$ = \Sexpr{maxcat$NH4} across all parameters in the category), \ac{chla} to \Sexpr{names(maxcat$Chla_mg_tot)} ($L1$ = \Sexpr{maxcat$Chla_mg_tot}), irradiance to \Sexpr{names(maxcat$irradiance)} ($L1$ = \Sexpr{maxcat$irradiance}), nitrate to \Sexpr{names(maxcat$NO3)} ($L1$ = \Sexpr{maxcat$NO3}), \ac{pom} to \Sexpr{names(maxcat$OM1)} ($L1$ = \Sexpr{maxcat$OM1}), \ac{dom} to \Sexpr{names(maxcat$OM2)} ($L1$ = \Sexpr{maxcat$OM2}), and phosphate to \Sexpr{names(maxcat$PO4)} ($L1$ = \Sexpr{maxcat$PO4}).  Finally, average sensititivy between parameter categories independent of the state variables ranged from \Sexpr{avecat[1]} for \Sexpr{names(avecat)[1]} (average $L1$ across all variables) to \Sexpr{avecat[length(avecat)]} for \Sexpr{names(avecat)[length(avecat)]}.

@@

<<eval = T>>=
data(cat_ident_all)
data(top_ident_all)

# percent combos less than gamma 15, 10 by state variable, by parameter category
# temperature category not evaluated, only one paramaeter
collcatpers <- lapply(cat_ident_all, function(stvar){

  lapply(stvar[!names(stvar) %in% 'Temperature'], function(x){

    colls <- x$coll
    l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
    l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
    c(l15, l10)

# percent combos less than gamma 15, 10 by state variable, by top in each category
colltoppers <- lapply(top_ident_all, function(stvar){

  lapply(stvar, function(x){

    colls <- x$coll
    l15 <- scinot(100 * sum(colls < 15)/length(colls), digits = 1)
    l10 <- scinot(100 * sum(colls < 10)/length(colls), digits = 1)
    c(l15, l10)

catpersO2 <- collcatpers$O2
toppersO2 <- colltoppers$O2
@

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Results for \ac{do} are provided first to demonstrate general concepts for the identifiability analyses, followed by an extension to the remaining state variables.  Parameter identifiability (as $\gamma$) for \ac{do} increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets (from two to many parameters) in the phytoplankton (\Sexpr{catpersO2$Phytoplankton[1]}\% less than $\gamma = 15$, \Sexpr{catpersO2$Phytoplankton[2]}\% less than $\gamma = 10$) and zooplankton categories (\Sexpr{catpersO2$Zooplankton[1]}\%, less than $\gamma = 15$, \Sexpr{catpersO2$Zooplankton[2]}\%, less than $\gamma = 10$), whereas a all combinations were identifiable for optics (\Sexpr{catpersO2$Optics[1]}\%, less than $\gamma = 15$, \Sexpr{catpersO2$Optics[2]}\%, less than $\gamma = 10$) and a majority identifiable for organic matter (\Sexpr{catpersO2$`Organic Matter`[1]}\%, less than $\gamma = 15$, \Sexpr{catpersO2$`Organic Matter`[2]}\%, less than $\gamma = 10$) combinations were identifiable. Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was \Sexpr{toppersO2$top1[1]}\% less than $\gamma = 15$ and \Sexpr{toppersO2$top1[2]}\% less than $\gamma = 10$ for the top parameters in each category, \Sexpr{toppersO2$top2[1]}\% and \Sexpr{toppersO2$top2[2]}\% for the top two, \Sexpr{toppersO2$top3[1]}\% and \Sexpr{toppersO2$top3[2]}\% for the top three, and \Sexpr{toppersO2$top4[1]}\% and \Sexpr{toppersO2$top4[2]}\% for the fop four.

An alternative view of the results in \cref{fig:identplo} to identify percent combinations that were identifiable as a function of number of parameters in each subset is show in

### 2016-11-30 ident_manu.tex
first paragraph of ident results done

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Tue Nov 29 17:13:44 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) evaluated parameters in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton), whereas average sensitivity among all parameters was $L1 = $ $0.01$. Within categories (excluding temperature), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $0.01$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton.  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $0$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All additional variables were sensitive to the same parameters (38 of 51 evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from $0.02$ for irradiance (\cref{tab:irrsens}) to $0.71$ for \ac{dom} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = $0.01$).  For each variable, $L1$ ranged from $2.24\times 10^{-6}$ (\textit{QminP}) to $8.49$ (\textit{mA}) for ammonium (\cref{tab:nh4sens}), $1.38\times 10^{-6}$ (\textit{QminP}) to $13.94$ (\textit{mA}) for \ac{chla} (\cref{tab:chlsens}), $1.92\times 10^{-7}$ (\textit{QminP}) to $0.13$ (\textit{ZKa}) for irradiance (\cref{tab:irrsens}), $6.67\times 10^{-7}$ (\textit{QminP}) to $8.49$ (\textit{umax}) for nitrate (\cref{tab:no3sens}), $6.41\times 10^{-5}$ (\textit{KNH4}) to $7.22$ (\textit{mA}) for \ac{pom} (\cref{tab:om1sens}),  $7.41\times 10^{-5}$ (\textit{KNH4}) to $14.25$ (\textit{mA}) for \ac{dom} (\cref{tab:om2sens}), and $8.21\times 10^{-7}$ (\textit{QminP}) to $1.47$ (\textit{ZKa}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to phytoplankton parameters (average $L1$ = $0.8$ across all parameters in the category), \ac{chla} to phytoplankton ($L1$ = ), irradiance to zooplankton ($L1$ = $0.03$), nitrate to zooplankton ($L1$ = $1.06$), \ac{pom} to temperature ($L1$ = $0.86$), \ac{dom} to temperature ($L1$ = $1.48$), and phosphate to zooplankton ($L1$ = $0.31$).  Finally, average sensititivy between parameter categories independent of the state variables ranged from $0.01$ for optics (average $L1$ across all variables) to $0.62$ for phytoplankton.

@@

%' data(cat_ident)
%' data(top_ident)
%' collcatpers <- lapply(cat_ident, function(x){
%' colltoppers <- lapply(top_ident, function(x){
%' maxtempcoll <- scinot(max(cat_ident$Temperature$coll))
%' The identifiability analyses suggested that most parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Parameter identifiability (as $\gamma$) increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identbox}).  By category, identifiability was lowest for parameter subsets in the phytoplankton (collcatpers$Phytoplankton[1]\% less than $\gamma = 15$, collcatpers$Phytoplankton[2]\% less than $\gamma = 10$)  and zooplankton categories (collcatpers$Zooplankton[1]\%, less than $\gamma = 15$, collcatpers$Zooplankton[2]\%, less than $\gamma = 10$), whereas a majority of combinations for temperature were identifiable (collcatpers$Temperature[1]\% less than $\gamma = 15$, collcatpers$Temperature[2]\%, less than $\gamma = 10$).  All subset combinations for optics and organic matter parameters had $\gamma < 10$.  All parameter subsets for choosing the top, top two, and top three parameters in each category were identifiable (\cref{fig:identbox}), whereas a majority were identifiable for choosing the top four (colltoppers$top4[1]\% less than $\gamma = 15$, colltoppers$top4[2]\% less than $\gamma = 10$) and top five (colltoppers$top4[1]\% less than $\gamma = 15$, colltoppers$top4[2]\%, less than $\gamma = 10$) parameters.  

>>>>

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Nov 30 14:30:44 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) evaluated parameters in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton), whereas average sensitivity among all parameters was $L1 = $ $0.01$. Within categories (excluding temperature), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $0.01$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $0$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All additional variables were sensitive to the same parameters (38 of 51 evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from $0.02$ for irradiance (\cref{tab:irrsens}) to $0.71$ for \ac{dom} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = $0.01$).  For each variable, $L1$ ranged from $2.24\times 10^{-6}$ (\textit{QminP}) to $8.49$ (\textit{mA}) for ammonium (\cref{tab:nh4sens}), $1.38\times 10^{-6}$ (\textit{QminP}) to $13.94$ (\textit{mA}) for \ac{chla} (\cref{tab:chlsens}), $1.92\times 10^{-7}$ (\textit{QminP}) to $0.13$ (\textit{ZKa}) for irradiance (\cref{tab:irrsens}), $6.67\times 10^{-7}$ (\textit{QminP}) to $8.49$ (\textit{umax}) for nitrate (\cref{tab:no3sens}), $6.41\times 10^{-5}$ (\textit{KNH4}) to $7.22$ (\textit{mA}) for \ac{pom} (\cref{tab:om1sens}),  $7.41\times 10^{-5}$ (\textit{KNH4}) to $14.25$ (\textit{mA}) for \ac{dom} (\cref{tab:om2sens}), and $8.21\times 10^{-7}$ (\textit{QminP}) to $1.47$ (\textit{ZKa}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to phytoplankton parameters (average $L1$ = $0.8$ across all parameters in the category), \ac{chla} to phytoplankton ($L1$ = $1.14$), irradiance to zooplankton ($L1$ = $0.03$), nitrate to zooplankton ($L1$ = $1.06$), \ac{pom} to temperature ($L1$ = $0.86$), \ac{dom} to temperature ($L1$ = $1.48$), and phosphate to zooplankton ($L1$ = $0.31$).  Finally, average sensititivy between parameter categories independent of the state variables ranged from $0.01$ for optics (average $L1$ across all variables) to $0.62$ for phytoplankton.

@@

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Results for \ac{do} are provided first to demonstrate general concepts for the identifiability analyses, followed by an extension to the remaining state variables.  Parameter identifiability (as $\gamma$) for \ac{do} increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets (from two to many parameters) in the phytoplankton ($60$\% less than $\gamma = 15$, $43$\% less than $\gamma = 10$) and zooplankton categories ($53.1$\%, less than $\gamma = 15$, $40$\%, less than $\gamma = 10$), whereas a all combinations were identifiable for optics ($100$\%, less than $\gamma = 15$, $100$\%, less than $\gamma = 10$) and a majority identifiable for organic matter ($91.9$\%, less than $\gamma = 15$, $76.5$\%, less than $\gamma = 10$) combinations were identifiable. Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was $100$\% less than $\gamma = 15$ and $100$\% less than $\gamma = 10$ for the top parameters in each category, $90.6$\% and $80.7$\% for the top two, $80.7$\% and $70.9$\% for the top three, and $55.8$\% and $45.7$\% for the fop four.

An alternative view of the results in \cref{fig:identplo} to identify percent combinations that were identifiable as a function of number of parameters in each subset is show in
